We are not the 99 percent: quantifying asphericity in the distribution of Local Group satellites

### 2018-02-15 paper/paper.tex
edit, remove repeated bibtex entries

\begin{abstract}
We use numerical simulations to build an explicit probability
distribution for the asphericity in the satellite distribution around
galaxies similar to the Local Group (LG) in the Lambda Cold Dark
Matter (LCDM) paradigm. 
This allows us to estimate the atipicallity
of the satellite distributions in the LG even when the underlying
simulations do not have enough systems that fully resemble the LG.
We demonstrate the method using three different simulations:
Illustris-1,  Illustris-1-Dark and ELVIS. 
Detailed results differ among the simulations suggesting a strong
influence of the typical Dark Matter (DM) halo mass in the different pair samples.  
However, there are three common trends in these simulations. 
First, at most of $2\%$ the pairs are expected to have satellite
distributions with the same asphericity as the LG; second,
between $27\%$ to $56\%$ of the pairs have a halo with a satellite
distribution as aspherical as in M31; and third, at most $3\%$ of the
pairs have a satellite distribution as planar as in the MW. 
The brightest satellites in the M31 galaxy turn out to have a rather
typical distribution in the LCDM context, while the MW distribution is atypical in
this respect.  
These results place the LG at the level of a $3\sigma$ outlier
in the LCDM paradigm. 
The approach presented here facilitates the comparison between
different numerical setups and choices to select halo pairs in simulations.
Quantifying the probability density for the satellite distribuition
asphericity as a function of halo mass and large scale environment 
will be a necessary step to understand the origin of the LG
atipicallity in the LCDM context as presented here.  
\end{abstract}

>>>>

\begin{abstract}
We use numerical simulations to build an explicit probability
distribution for the asphericity in the satellite distribution around
galaxies similar to the Local Group (LG) in the Lambda Cold Dark
Matter (LCDM) paradigm. 
This allows us to estimate the atipicallity
of the satellite distributions in the LG even when the underlying
simulations do not have enough systems that fully resemble the LG.
We demonstrate the method using three different simulations:
Illustris-1,  Illustris-1-Dark and ELVIS. 
Detailed results differ among the simulations suggesting a strong
influence of the typical Dark Matter (DM) halo mass in the different
pair samples. 
However, there are three common trends in these simulations. 
First, at most of $2\%$ the pairs are expected to have satellite
distributions with the same asphericity as the LG; second,
between $27\%$ to $56\%$ of the pairs have a halo with a satellite
distribution as aspherical as in M31; and third, at most $3\%$ of the
pairs have a satellite distribution as planar as in the MW. 
The brightest satellites in the M31 galaxy turn out to have a rather
typical distribution in the LCDM context, while the MW distribution is atypical in
this respect.  
These results place the LG at the level of a $3\sigma$ outlier in the
LCDM paradigm. 
Understand this atipicallity will require quantifying the probability
distribution for the satellite distribuition asphericity as a function
of halo mass and large scale environment. 
The approach presented can facilitate this study and other comparisons
between different numerical setups and choices to select LG pairs in
simulations.  
\end{abstract}
